Classless Java: Tuning Java Interfaces

### 2015-11-25 paper/sections/CaseStudy.tex
EP case study revised

The \emph{Expression Problem} (EP)~\cite{wadler98expression}
is a well-known problem about modular extensibility issues in software
evolution. Recently, a new solution~\cite{eptrivially} that uses only
covariant type refinement that has been shown to work in Java. When
that solution is modelled with interfaces and default methods, it can
even provide independent extensibility: that is, the ability to
assemble a system, from multiple, independently developed extensions.
Unfortunatelly, a plain Java solution has alot of code verbosity in
Java due to code required for instantiation, which makes the solution 
cumbersome to use. The \mixin annotation helps to remove such code, and 
makes the solution quite easy to use.
\bruno{We should try to do some space saving in here. You could, for example 
combine the code for the initial system with the code for a new operation in a figure/table 
with 2 colums. Then just refer to the Table/Figure in the text.}

@@

\lstinputlisting[linerange=34-44]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_INIT

@@

\lstinputlisting[linerange=48-53]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_SUB

@@

\lstinputlisting[linerange=57-66]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_PRINT
\bruno{Please add a Java comment to the code emphasizing that the types of the getters are refined!}
\bruno{please indent code properly (the getters are not aligned with default))}

@@

\paragraph{Independent Extensibility}
To show that our approach supports independent extensibility~\cite{zenger05independentlyextensible}, we first define a
new operation \texttt{collectLit} (which collects all
literal components in an expression) on expressions. For space reasons, 
we omit the definitions of the methods: 
\bruno{No need to show full code in the paper. You can just put \ldots in the body of the collectLit methods.}

\lstinputlisting[linerange=70-86]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_COLLECTLIT

@@

\lstinputlisting[linerange=90-94]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=INDEPENDENT_EXTENSIBILITY

@@

Note that the code for instantiation is automatically generated by \mixin. 
So, for example creating a simple expression of type \texttt{ExpPC} is 
as simple as:

\bruno{Add code here}

\noindent In contrast, in a pure Java solution, the tedious instantiation code would need 
to be defined manually.

>>>>

The \emph{Expression Problem} (EP)~\cite{wadler98expression}
is a well-known problem about modular extensibility issues in software
evolution. Recently, a new solution~\cite{eptrivially} that uses only
covariant type refinement that has been shown to work in Java. When
that solution is modelled with interfaces and default methods, it can
even provide independent extensibility: that is, the ability to
assemble a system, from multiple, independently developed extensions.
Unfortunatelly, a plain Java solution has alot of code verbosity in
Java due to code required for instantiation, which makes the solution 
cumbersome to use. The \mixin annotation helps to remove such code, and 
makes the solution quite easy to use.
% \bruno{We should try to do some space saving in here. You could, for example 
% combine the code for the initial system with the code for a new operation in a figure/table 
% with 2 colums. Then just refer to the Table/Figure in the text.}

@@

\begin{lstlisting}[multicols=2]
interface Exp { int eval(); }
@Obj interface Lit extends Exp {
	int x();
	default int eval() { return x(); }
}
@Obj interface Add extends Exp {
	Exp e1(); Exp e2();
	default int eval() {
		return e1().eval() + e2().eval();
	}
}
\end{lstlisting}

% \begin{multicols}{2}
% \lstinputlisting[linerange=34-44]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_INIT
% \end{multicols}

@@

\lstinputlisting[linerange=48-49]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_SUB

@@

\lstinputlisting[linerange=53-61]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_PRINT
%\bruno{Please add a Java comment to the code emphasizing that the types of the getters are refined!}
%\bruno{please indent code properly (the getters are not aligned with default))}

@@

\paragraph{Independent Extensibility}
To show that our approach supports independent extensibility~\cite{zenger05independentlyextensible}, we first define a
new operation \texttt{collectLit} (which collects all
literal components in an expression) on expressions. For space reasons, 
we omit the definitions of the methods:

\begin{lstlisting}[]
interface ExpC extends Exp { List<Integer> collectLit(); }
@Obj interface LitC extends Lit, ExpC {
    default List<Integer> collectLit() { ... }}
@Obj interface AddC extends Add, ExpC {
    ExpC e1(); ExpC e2(); //return type refined!
    default List<Integer> collectLit() { ... }}
\end{lstlisting}

@@

\lstinputlisting[linerange=85-87]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=INDEPENDENT_EXTENSIBILITY

@@

Note that the code for instantiation is automatically generated by \mixin. 
So, for example creating a simple expression of type \texttt{ExpPC} is 
as simple as:
\begin{lstlisting}
ExpPC e8 = AddPC.of(LitPC.of(3), LitPC.of(4));
\end{lstlisting}
\noindent In contrast, in a pure Java solution, the tedious instantiation code would need 
to be defined manually.
